---
title: "Data preparation and cleaning"
output: html_notebook
---
# Packages
```{r}
pacman::p_load(tidyverse,
               kableExtra, 
               here, 
               janitor,
               MatchIt,
               scales, 
               # epiR, 
                table1, # contains the label function
               gtsummary, 
               # tableone,
               # survival, 
               ggpubr, 
               RColorBrewer, 
               # ggmap,
               # geojsonR,
                leaflet,
               # rgdal,
              #  sf,
               viridis)
```

# Dataset
Hidden for anonimity
```{r read csv, include=FALSE}
df <- read_csv("Dati_toothpaste_latvia_lithuania.csv") # data until march 14, 2019
```


```{r}
theme_set(theme_minimal())
```


```{r}
df <- janitor::clean_names(df)
```

# Data cleaning
```{r visdat}
# visdat::vis_dat(df)
```

```{r}
names(df)
```

Clean names and remove empty cols


Age groups
```{r create age groups}
df <- df %>% 
  mutate(age_group = case_when(
    `age` < 6 ~ "0 ~ 5", 
    `age` < 13 ~ "6 ~ 12", 
    `age` < 19 ~ "13 ~ 18", 
    TRUE ~ "> 18"
  ))


#  relevel age groups
df$age_group <- fct_relevel(df$age_group, "0 ~ 5" , "6 ~ 12", "13 ~ 18", "> 18" )

```

```{r}
janitor::tabyl(df$age_group) %>% janitor::adorn_pct_formatting() # # table age groups
```


IN age group change NA to adult

```{r}
df$`children_adult` <- df$`children_adult` %>% 
  replace_na("adult")
```

## Grouping fluoride concentrations
Create a new variable F_concentration_group

First, convert the Not known in some nomber to recod afterwards

```{r}
df <- df %>% 
  mutate(f_concentration = fct_recode(f_concentration, "99999" = "Not known")) 
```


```{r F groups}
df <- df %>%
  # some pre processing
  mutate(f_concentration = as.character(f_concentration)) %>%  # was a factor, so first convert to character
  mutate(f_concentration = as.integer(f_concentration)) %>% # f_concentration was stored as character and not as integer!
  
  # create the new categories
  mutate(
    `f_concentration_group` = case_when(
      `f_concentration` == 99999 ~ "Not known",
      `f_concentration` < 1 ~ "No fluoride",
      `f_concentration` < 1000 ~ "<1000 ppm",
      `f_concentration` < 1400 ~ "1000-1399 ppm",
      `f_concentration` < 1501 ~ "1400-1500 ppm",
      TRUE ~ "NA"
    )
  ) %>%
  # check the table
  # tabyl(f_concentration_group)
  
  # now reorder
  mutate(
    `f_concentration_group` = fct_relevel(
      f_concentration_group,
      "Not known",
      "No fluoride",
      "<1000 ppm",
      "1000-1399 ppm",
      "1400-1500 ppm"
    )
  ) 
```

```{r}
# check
df %>% 
  tabyl(f_concentration_group, country)
```

ok, first change the NA to other values, as Not Available

```{r}

df <- df %>% 
  mutate(f_concentration_group = fct_recode(f_concentration_group, "Not available" = "NA")) 

```



## Data imputation

```{r}
pacman::p_load(mice, # the package for imputation
               naniar) # to visualize the NAs
```


```{r}
df <- replace_with_na_at(data = df,
                         .vars = "f_concentration_group",
                         condition = ~.x == "Not known") %>% 
  droplevels()
```


```{r}
df %>% 
  tabyl(f_concentration_group, country)
```

check the missing pattern
```{r}
df %>% 
  select(country, f_concentration_group) %>% 
  md.pattern()
```
```{r}
df %>% 
  select(country, f_concentration_group) %>% 
  gg_miss_var(country) 
```



```{r}
# Store a copy of the previous dataset
# df_non_imputed <- df
```

```{r}
df %>% 
  tabyl(f_concentration_group, country) %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns()
```

```{r}
# Define the levels
levels <- c("<1000 ppm", "1000-1399 ppm", "1400-1500 ppm", "No fluoride", "Not available")
num_na <- c(28, 36, 146, 39, 6)

# Create the seq
seq <- rep(levels, num_na)

# Order the DF
df <- df[order(is.na(df$f_concentration_group)), ]

# Assign the values
df$f_concentration_group[is.na(df$f_concentration_group)] <- seq
```


check

```{r}
df %>% 
  tabyl(f_concentration_group, country)
```



```{r}
write_csv(df, "df.csv")
```

```{r}
rm(df)
```
