---
title: "2019 Toothpaste Latvia"
output: 
  html_notebook: 
    toc: yes
    toc_float: true
    fig_caption: true
---
# Packages
```{r, warning=FALSE, results='hide'}
library(tidyverse)
if (!require("visdat")) install.packages("visdat") #instala paquetes en caso que no est?n instalados
if (!require("kableExtra")) install.packages("kableExtra") #instala paquetes en caso que no est?n instalados
```

# Dataset
Hidden for anonimity
```{r read csv, include=FALSE}
# df <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vS_CyX5w7gOyekyzKDiwO2LQ9h5q9r5CLgNISIFXfED3X7IXNAEES5Kg_4vCyK_Q10UDCt_RZtLYyO7/pub?output=csv")
df <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vQo5nCKqjmclsLAUdGF3spDrwRAIqLT1ChCDj8yQ6DcgO0CDwhdQRIw3lMbA9du1FEdaR084FI-Xknc/pub?gid=861337044&single=true&output=csv") # data until march 14, 2019
```
# Data cleaning
```{r visdat}
visdat::vis_dat(df)
```

```{r}
names(df)
```

Clean names and remove empty cols
```{r clean names and remove columns}
df <- df %>% 
  # janitor::clean_names(.) %>% 
  janitor::remove_empty(., c("rows", "cols")) %>% 
  select(-c(`F concentration for the second paste`, `Laika zīmogs`))
```

Age groups
```{r create age groups}
df <- df %>% 
  mutate(age_group = case_when(
    `Age` < 6 ~ "0 ~ 5", 
    `Age` < 13 ~ "6 ~ 12", 
    `Age` < 19 ~ "13 ~ 18", 
    TRUE ~ "> 18"
  ))


#  relevel age groups
df$age_group <- fct_relevel(df$age_group, "0 ~ 5" , "6 ~ 12", "13 ~ 18", "> 18" )

# table age groups
table(df$age_group)
```


IN age group change NA to adult

```{r}
df$`Children/Adult` <- df$`Children/Adult` %>% replace_na("adult")
```

# EDA
```{r check names}
names(df)
```
```{r summary}
summary(df)
```
Convert F to number
```{r Convert F to number}
df$`F concentration 2` <- as.integer(df$`F concentration`)
```
```{r Check distribution of F in toothpaste}
df %>% 
  ggplot(aes(x = `F concentration 2`)) + 
  geom_histogram(bins = 5)
```
Create a new var grouping F concentration
```{r F groups}
df <- df %>% 
  mutate(`f_concentration` = case_when(
    `F concentration 2` < 1 ~ "No fluoride", 
    `F concentration 2` < 550 ~ "1 to 550 ppm", 
    `F concentration 2` < 1000 ~ "551 to 999 ppm",
    `F concentration 2` < 1501 ~ "1000 and 1500 ppm", 
    `F concentration 2` > 1501 ~ ">1500 ppm", 
    TRUE ~ "NA"
  ))
```

```{r table concentration}
table(df$f_concentration)
```

```{r}
df$f_concentration <- factor(df$f_concentration,, 
                             levels = c(
                              "NA",
                              "No fluoride", 
                              "1 to 550 ppm", 
                              "551 to 999 ppm", 
                              "1000 and 1500 ppm"))
```





Check as table
```{r F Main table}
prop.table((table(df$`f_concentration`))) * 100
```
Check with graph
```{r F main table as graph}
df %>% 
  # na.omit() %>% 
  ggplot(aes(x = `f_concentration`)) + 
  geom_bar() + 
  coord_flip() + 
  theme_minimal()
```
Check groups of F concentration by age
```{r F by age boxplot}

df %>% 
  # na.omit() %>% 
  ggplot(aes(x = `f_concentration`, y = `Age`)) +
  geom_boxplot() +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "Age distribution per toothpaste F concentration type", 
       x = "Fluoride concentration", 
       y = "Age (years)") 

```


```{r plot porcentaje por edad}
df %>% 
  na.omit(f_concentration) %>% 
  group_by(age_group, f_concentration) %>% 
  summarise(count = n()) %>% 
  mutate(perc=count/sum(count)) %>% 
  ggplot(aes(x = age_group, y = perc*100, fill = f_concentration)) +
  geom_bar(stat="identity") + 
  facet_grid(f_concentration ~ .) + 
  theme_minimal() + 
  labs(title = "Fluoride Concentration toothpaste by age group", 
       y = "Percentage", 
       x = "Age group", 
       fill = "Fluoride concentration")

```
```{r}
df %>% 
  na.omit(f_concentration) %>% 
  group_by(age_group, f_concentration) %>% 
  summarise(count = n()) %>% 
  mutate(perc=count/sum(count)) %>% 
  ggplot(aes(x = age_group, y = perc*100, fill = f_concentration)) +
  geom_bar(stat="identity") + 
  facet_wrap(~f_concentration) + 
  theme_minimal() + 
  labs(title = "Fluoride Concentration toothpaste by age group", 
       y = "Percentage", 
       x = "Age group", 
       fill = "Fluoride concentration")
```
```{r}
df %>% 
  na.omit(f_concentration) %>% 
  group_by(age_group, f_concentration) %>% 
  summarise(count = n()) %>% 
  mutate(perc=count/sum(count)) %>% 
  ggplot(aes(x = age_group, y = perc*100, fill = f_concentration)) +
  geom_bar(stat="identity") + 
  facet_wrap(~f_concentration) + 
  theme_minimal() + 
  labs(title = "Fluoride Concentration toothpaste by age group", 
       y = "Percentage", 
       x = "Age group", 
       fill = "Fluoride concentration") + 
  coord_flip()
```



Table f concentration by age with means, etc
```{r table F by group age mean sd etc}
df %>% 
  group_by(`f_concentration`) %>% 
  summarise(n = n(), 
            mean = mean(`Age`), 
            sd = sd(`Age`), 
            min = min(`Age`), 
            max = max(`Age`))
```
```{r plot by city}
df %>% 
  na.omit(f_concentration) %>% 
  group_by(`Lives in`, f_concentration) %>% 
  summarise(count = n()) %>% 
  mutate(perc=count/sum(count)) %>% 
  ggplot(aes(x = `Lives in`, y = perc*100, fill = f_concentration)) +
  geom_bar(stat="identity") + 
  facet_wrap(~f_concentration) + 
  theme_minimal() + 
  labs(title = "Fluoride Concentration toothpaste by location", 
       y = "Percentage", 
       x = "Location", 
       fill = "Fluoride concentration")
```
```{r plot by city 2}
df %>% 
  na.omit(f_concentration) %>% 
  group_by(`Lives in`, f_concentration) %>% 
  summarise(count = n()) %>% 
  mutate(perc=count/sum(count)) %>% 
  ggplot(aes(x = `Lives in`, y = perc*100, fill = f_concentration)) +
  geom_bar(stat="identity") + 
  facet_wrap(~f_concentration) + 
  theme_minimal() + 
  labs(title = "Fluoride Concentration toothpaste by location", 
       y = "Percentage", 
       x = "Location", 
       fill = "Fluoride concentration") + 
  coord_flip()
```

Cuantos nombres de pastas hay?
```{r check toothpastes names}
df %>% 
  group_by(Toothpaste) %>% 
  summarise(count = n())
```




```{r table age by concentration}
df %>% 
  janitor::tabyl(age_group, `f_concentration`) %>% 
  knitr::kable()
```


```{r table live by concentration}
df %>% 
  janitor::tabyl(`Lives in`, `f_concentration`) %>% 
  janitor::adorn_totals(c("row", "col")) %>%
  janitor::adorn_percentages("row") %>% 
  janitor::adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  janitor::adorn_ns() %>%
  janitor::adorn_title("combined") %>%
  knitr::kable()
```


```{r nice table age by concentration}
df %>% 
  janitor::tabyl(age_group, `f_concentration`) %>% 
  janitor::adorn_totals(c("row", "col")) %>%
  janitor::adorn_percentages("row") %>% 
  janitor::adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  janitor::adorn_ns() %>%
  janitor::adorn_title("combined") %>%
  knitr::kable() %>% kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```



```{r see names F free toothpastes }
df %>%  
  filter(`f_concentration` == "0 - No fluoride") %>% 
  #na.omit() %>%    # Using "data", filter out all rows with NAs in aa 
  group_by(`Toothpaste`) %>%          # Then, with the filtered data, group it by "bb"
  summarise(Unique_Elements = n_distinct(`Toothpaste`))  %>%   # Now summarise with unique elements per group
  knitr::kable()
```

See all toothpastes


```{r see names all toothpastes }
df %>%  
  filter(!`f_concentration` == "0 - No fluoride") %>% 
  #na.omit() %>%    # Using "data", filter out all rows with NAs in aa 
  group_by(`Toothpaste`) %>%          # Then, with the filtered data, group it by "bb"
  summarise(Unique_Elements = n_distinct(`Toothpaste`))  %>%   # Now summarise with unique elements per group
  knitr::kable()
```

# SIGNIFICANCE TEST

## Differences per region?

```{r}

table(df$`Lives in`, df$f_concentration)
```
```{r}
chisq.test(table(df$`Lives in`, df$f_concentration))
```
```{r}
mosaicplot(table(df$f_concentration, df$`Lives in`), shade = T)
```



## Differences per sex?

```{r}
chisq.test(table(df$age_group, df$f_concentration))
```
```{r}
mosaicplot(table(df$f_concentration, df$age_group), shade = T)
```


# SAMPLE SIZE CALCULATIONS

```{r add a new var to calculate variance}
df <- df %>% 
  mutate(
    F_variance = case_when(
      `F concentration` < "1450" ~ "0", 
      TRUE ~ "1"
    )
  )
```

Calculate var intra and inter
Intra
```{r var intra}
intra <- var(df$F_variance)
intra
```
Inter
```{r}
inter <- df %>% 
  group_by(`School code`) %>% 
  summarise(variance = var(F_variance)) %>% 
  summarise(variance_inter = mean(variance))
inter
```
```{r}
inter / (intra + inter)
```
f the ICC is known, for example from a pilot study, it can be used at the design stage of the trial to inform the sample size calculation. The ‘design effect’ (DE) can be used to estimate the extent to which the sample size should be inflated to accommodate for the homogeneity in the clustered data:

      DE = 1+(n-1)ρ

      n = average cluster size(      ρ = ICC for the desired outcome

The DE can then be used to calculate the ‘effective sample size’. This is the ‘real’ sample size in a clustered trial, compared with the number of participants actually enrolled in the study. It is calculated using the formulae below:
mk / DE, where m = unites per cluster, k = clusters

Hence





## Variance inter and intra for rho
```{r variance intra}
en <- df %>% 
  na.omit() %>% 
  summarise(variance = var(`F concentration`))
en
```

```{r variance inter}
entre <- df %>% 
  na.omit() %>% 
  group_by(`School code`) %>% 
  summarise(variance = var(`F concentration`)) %>% 
  summarise(variance_mean = mean(variance))
entre
```

```{r icc}
new_rho <- entre/(entre + en)
```

## Another variance


```{r, warning=FALSE, message=FALSE}
install.packages("epiR")
library(epiR)
```
## Data

```{r}
p <- 0.50 # expected proportion people using non-fluoridated toothpaste
b <- 25 # units to be examined per school (cluster)
```

## EXAMPLE 1
 The expected prevalence of disease in the population  is `r p`.
 We wish to conduct a survey, sampling `r b` persons per school. No data
 are available to provide an estimate of rho, though we suspect
 the intra-cluster correlation for this disease to be moderate.
 We wish to be 95% certain of being within 10% of the true population
 prevalence of disease. How many schools should be sampled?

```{r}

D <- 4
rho <- (D - 1) / (b - 1)
cluster_ss <- epiR::epi.clustersize(p = p, # expected prevalence
                b = b, # sampling per school
                rho = rho, # difference among variance inside the clusters vs between clusters, i.e. caries prevalence inside every school expected to be homogeneous, and prevalence between schools expected to be homogenous also. 
                epsilon.r = 0.2, #the acceptable relative error. Relative error is error / expected prevalence
                conf.level = 0.95)

```

Needed schools 
```{r need schools 1}
cluster_ss$clusters
```
Sample size total
```{r total sample size 1}
cluster_ss$units
```


**We need to sample `r cluster_ss$clusters` schools (`r cluster_ss$units` samples in total)**
 
Reference: 
Otte J, Gumm I (1997). Intra-cluster correlation coefficients of 20 infections calculated from the results of cluster-sample surveys. Preventive Veterinary Medicine 31: 147 - 150.


## EXAMPLE 2
 A cross-sectional study is to be carried out to determine the prevalence
 of a given disease in a population using a two-stage cluster design. We
 estimate prevalence to be `r p` and we expect rho to be in the order of 0.02.
 We want to take sufficient samples to be 95% certain that our estimate of
 prevalence is within 5% of the true population value (that is, a relative
 error of 0.05 / 0.50 = 0.1). Assuming `r b` responses from each cluster,
 how many clusters do we need to be sample?


```{r}
cluster_ss_2 <- epiR::epi.clustersize(p = p, # expected prevalence
                b = b, # participants per cluster
                rho = 0.02, # with new_rho is crazy
                epsilon.r = 0.1,
                conf.level = 0.95)
```
Schools 2
```{r schools 2}
cluster_ss_2$clusters
```
Sample 2
```{r sample 2}
cluster_ss_2$units
```

**We need to sample `r cluster_ss_2$clusters` clusters (`r cluster_ss_2$units` samples in total)**

Reference: 
Bennett S, Woods T, Liyanage WM, Smith DL (1991). A simplified general method for clustersample
surveys of health in developing countries. World Health Statistics Quarterly 44: 98 - 106.


# Sampling procedure (if required)
```{r, warning=FALSE}
# library(sampling)
```

 Two-stage cluster sampling
 Uses the 'swissmunicipalities' data 
 the variable 'REG' (region) has 7 categories;
 it is used as clustering variable in the first-stage sample
 the variable 'CT' (canton) has 26 categories; 
 it is used as clustering variable in the second-stage sample
 4 clusters (regions) are selected in the first-stage 
 1 canton is selected in the second-stage from each sampled region 
 the method is simple random sampling without replacement in each stage
 (equal probability, without replacement)



data(swissmunicipalities)
c = swissmunicipalities
c = c[order(c$REG,c$CT),]
attach(c)

m = mstage(c,
           stage = list("cluster","cluster"),
           varnames = list("REG","CT"),
           size = list(4,c(1,1,1,1)),
           method = list("srswor","srswor"))

 the first stage is m[[1]], the second stage is m[[2]]
 the selected regions

unique(m[[1]]$REG)

the selected cantons

unique(m[[2]]$CT)

extracts the observed data

x = getdata(c,m)[[2]]

check the output

table(x$REG,x$CT)

# PENDIENTE

Cuantas familias tienes pastas de niños?
df %>% 
  group_by(`Children/Adult`, `Family code`) %>% 
  summarise(n = n()) %>% 
  spread(`Children/Adult`, n)

