---
title: "2019 Toothpaste Latvia"
date-modified: last-modified
format:
  html:
    toc: true
toc-expand: 3
code-fold: true
code-tools: true
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

# Packages

```{r}
pacman::p_load(tidyverse,
               kableExtra, 
               here, 
               janitor,
               MatchIt,
               scales, 
               # epiR, 
                table1, # contains the label function
               gtsummary, 
               # tableone,
               # survival, 
               ggpubr, 
               RColorBrewer, 
               patchwork, # to combine plots
               # ggmap,
               # geojsonR,
                leaflet,
               # rgdal,
              #  sf,
               viridis)
```

# Dataset

Hidden for anonimity

```{r read csv, include=FALSE}
df <- read_csv("df.csv") # data until march 14, 2019
```

```{r}
theme_set(theme_minimal())
```

```{r}
# reorder the levels of f_concentration_group

df <- df |> 
  mutate(f_concentration_group = fct_relevel(f_concentration_group, 
                                             "Not available", 
                                            "No fluoride", 
                                            "<1000 ppm", 
                                            "1000-1399 ppm", 
                                            "1400-1500 ppm"
                                            ))
```

# Separate datasets for each country

```{r}
df_latvia <- df |> 
  filter(country == "Latvia")
```

```{r}
df_lithuania <- df |> 
  filter(country == "Lithuania")
```

# EDA

```{r check names}
# names(df)
```

```{r summary df}
summary(df) |> 
  kable() |> 
  kable_classic(full_width = F)
```

```{r summary LV}
summary(df_latvia) |> 
  kable() |> 
  kable_classic(full_width = F)
```

```{r summary LT}
summary(df_lithuania)|> 
  kable() |> 
  kable_classic(full_width = F)
```

Convert F to number

```{r Convert F to number}
df$`f_concentration` <- as.integer(df$`f_concentration`)
```

```{r Check distribution of F in toothpaste}
# df |> 
#   ggplot(aes(x = `f_concentration`)) + 
#   geom_histogram(bins = 5)
```

Convert F2 to number

```{r Convert F2 to number}
# df$`f_concentration_2` <- as.integer(df$`f_concentration_for_the_second_paste`)
```

```{r Check distribution of F2 in toothpaste}
# df |> 
#  ggplot(aes(x = `f_concentration_for_the_second_paste`)) + 
#  geom_histogram(bins = 5)
```

# Table 1

```{r}
df$children_adult <-
  factor(df$children_adult,
         labels = c("Adult toothpaste",
                    "Child toothpaste"))

label(df$age) <- "Age"
label(df$age_group) <- "Age groups"
label(df$lives_in) <- "Lives In"
label(df$lives_in_specific) <- "Region"
label(df$f_concentration_group) <- "F concentration in toothpaste"
label(df$nr_of_family_members) <- "Number of family members"




```

Relevel age groups

```{r}
df <- df |> 
  mutate(age_group = fct_relevel(age_group, 
                                 "0 ~ 5", "6 ~ 12", "13 ~ 18", "> 18"
                                 ))
```

```{r}
df |> 
  select(age, age_group, lives_in, country) |> 
  gtsummary::tbl_summary(by = country, missing = "no", 
                         statistic = list(all_continuous() ~ "{mean} ({sd})")) |> 
  gtsummary::add_overall()
```

# Table of main results

```{r}
df |> 
  select(f_concentration_group, children_adult, country) |> 
  gtsummary::tbl_summary(by = country, missing = "no", 
                         statistic = list(all_continuous() ~ "{mean} ({sd})")) |> 
  gtsummary::add_overall() |> 
  gtsummary::add_p()
```

Difference by country

```{r}
# Perform chi-squared test
chisq.test(df$country, df$f_concentration_group) 
```

Difference by group

```{r}
# Perform chi-squared test
chisq.test(df$children_adult, df$f_concentration_group)
```

```{r}
mosaicplot(table(df$country, df$f_concentration_group), shade = T)
```

```{r}
df_withoutNA <- df |>
  filter(f_concentration_group != "NA")
```

```{r}
mosaicplot(table(df_withoutNA$country, df_withoutNA$f_concentration_group), shade = T)
```

```{r}

contingency_tables_list <- table(df$f_concentration_group, df$country, df$children_adult)
contingency_tables_list 


```

```{r}
# Perform the Mantel-Haenszel test
mantel_haenszel_test <- mantelhaen.test(contingency_tables_list)

# Print the results
mantel_haenszel_test

```

```{r}
rm(contingency_tables_list, mantel_haenszel_test, df_withoutNA)
```

```{r Table 1_1}
df |> 
  select(lives_in, f_concentration_group) |> 
  gtsummary::tbl_summary(by = lives_in, missing = "no") |> 
  gtsummary::add_overall()
```

## Table of main results for Latvia

```{r Table 1_2}
df |> 
  filter(country == "Latvia") |> 
  select(f_concentration_group, age_group) |> 
  gtsummary::tbl_summary(by = age_group, missing = "no") |> 
  gtsummary::add_overall()
```

```{r Table 1_3}
df |> 
  filter(country == "Latvia") |> 
  select(lives_in, f_concentration_group) |> 
  gtsummary::tbl_summary(by = lives_in, missing = "no") |> 
  gtsummary::add_overall()
```

## Table of main results for Lithuania

```{r}
df |> 
  filter(country == "Lithuania") |> 
  select(f_concentration_group, age_group) |> 
  gtsummary::tbl_summary(by = age_group, missing = "no") |> 
  gtsummary::add_overall()
```

```{r Table 1}
df |> 
  filter(country == "Lithuania") |> 
  select(lives_in, f_concentration_group) |> 
  gtsummary::tbl_summary(by = lives_in, missing = "no") |> 
  gtsummary::add_overall()
```

# Graph F concentration by age groups

## Figure 2 final

```{r}
my_palette <- brewer.pal(name="Spectral", n = 7)[0:9]
```

```{r}
latvia_age_plot <- df |>
  select(age_group, country, f_concentration_group) |>
  filter(f_concentration_group != "NA") |>
  filter(country == "Latvia") |> 
  # calculate the counts
  group_by(country, f_concentration_group, age_group) |>
  count() |>
  
  # calculate the total counts for each age_group
  group_by(country, age_group) |>
  mutate(total_count = sum(n)) |>
  
  # calculate the percentages
  ungroup() |>
  mutate(percentage = (n / total_count) * 100) |> 
  
  # select the columns you want to keep
  select(country, f_concentration_group, age_group, n, percentage) |>
  # pivot wider
  select(f_concentration_group, age_group, percentage) |>
  # the plot
  ggplot(aes(
    x = age_group,
    label = percentage,
    y = percentage,
    fill = f_concentration_group
  )) +
  geom_col() +
  # adjust the scale
  scale_y_continuous(labels = label_percent(scale = 1)) +
  # change the colors
  scale_fill_brewer(palette = "Spectral") +
  # the labels
geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5), size = 3) +
  labs(title = "Use of F Toothpaste by Age", 
       subtitle = "Latvia", 
       x = "Age group", 
       y = "", 
       fill = "") + 
  theme(legend.position = "none")
```

## 

```{r}
lithuania_age_plot <- df |>
  select(age_group, country, f_concentration_group) |>
  filter(f_concentration_group != "NA") |>
  filter(country == "Lithuania") |> 
  # calculate the counts
  group_by(country, f_concentration_group, age_group) |>
  count() |>
  
  # calculate the total counts for each age_group
  group_by(country, age_group) |>
  mutate(total_count = sum(n)) |>
  
  # calculate the percentages
  ungroup() |>
  mutate(percentage = (n / total_count) * 100) |> 
  
  # select the columns you want to keep
  select(country, f_concentration_group, age_group, n, percentage) |>
  # pivot wider
  select(f_concentration_group, age_group, percentage) |>
  # the plot
  ggplot(aes(
    x = age_group,
    label = percentage,
    y = percentage,
    fill = f_concentration_group
  )) +
  geom_col() +
  # adjust the scale
  scale_y_continuous(labels = label_percent(scale = 1)) +
  # change the colors
  scale_fill_brewer(palette = "Spectral") +
  # the labels
geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5), size = 3) +
  labs(subtitle = "Lithuania", 
       x = "Age group", 
       y = "", 
       fill = "F Concentration Group", 
       caption = "Note: Due to rounding, the sum of percentages may not equal 100.")
```

```{r}
latvia_age_plot / lithuania_age_plot
  
```

```{r}
ggsave(
  here("figures", "figure_age.pdf"),
  width = 20, height = 20, units = "cm", 
  # res = 300, 
  device='pdf'
)
```

```{r}
rm(latvia_age_plot, lithuania_age_plot)
```

## Differences per age group

```{r}
table(df$age_group, df$f_concentration_group) |> 
  kable() |> 
  kable_classic(full_width = F)
```

```{r}
chisq.test(table(df$age_group, df$f_concentration_group))
```

```{r}
mosaicplot(table(df$f_concentration_group, df$age_group), shade = T)
```

## Latvia per age

```{r}

table(df_latvia$age_group, df_latvia$f_concentration_group)|> 
  kable() |> 
  kable_classic(full_width = F)
```

```{r}
chisq.test(table(df_latvia$age_group, df_latvia$f_concentration_group))
```

```{r}
mosaicplot(table(df_latvia$f_concentration_group, df_latvia$age_group), shade = T)
```

## Lithuania per age

```{r}
  table(df_lithuania$age_group, df_lithuania$f_concentration_group)|> 
  kable() |> 
  kable_classic(full_width = F)
```

```{r}
chisq.test(table(df_lithuania$age_group, df_lithuania$f_concentration_group))
```

```{r}
mosaicplot(table(df_lithuania$f_concentration_group, df_lithuania$age_group), shade = T)
```

# Graph F concentration by living area

## Figure 3 final

```{r}
latvia_Living_area_plot <- df |>
  select(lives_in, country, f_concentration_group) |>
  filter(f_concentration_group != "NA") |>
  filter(country == "Latvia") |> 
  # calculate the counts
  group_by(country, f_concentration_group, lives_in) |>
  count() |>
  
  # calculate the total counts for each lives_in
  group_by(country, lives_in) |>
  mutate(total_count = sum(n)) |>
  
  # calculate the percentages
  ungroup() |>
  mutate(percentage = (n / total_count) * 100) |> 
  
  # select the columns you want to keep
  select(country, f_concentration_group, lives_in, n, percentage) |>
  # pivot wider
  select(f_concentration_group, lives_in, percentage) |>
  # the plot
  ggplot(aes(
    x = lives_in,
    label = percentage,
    y = percentage,
    fill = f_concentration_group
  )) +
  geom_col() +
  # adjust the scale
  scale_y_continuous(labels = label_percent(scale = 1)) +
  # change the colors
  scale_fill_brewer(palette = "Spectral") +
  # the labels
geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5), size = 3) +
  labs(title = "Use of F Toothpaste by Living Area", 
       subtitle = "Latvia", 
       x = "Living Area group", 
       y = "", 
       fill = "") + 
  theme(legend.position = "none")
```

```{r}
lithuania_Living_area_plot <- df |>
  select(lives_in, country, f_concentration_group) |>
  filter(f_concentration_group != "NA") |>
  filter(country == "Lithuania") |> 
  # calculate the counts
  group_by(country, f_concentration_group, lives_in) |>
  count() |>
  
  # calculate the total counts for each lives_in
  group_by(country, lives_in) |>
  mutate(total_count = sum(n)) |>
  
  # calculate the percentages
  ungroup() |>
  mutate(percentage = (n / total_count) * 100) |> 
  
  # select the columns you want to keep
  select(country, f_concentration_group, lives_in, n, percentage) |>
  # pivot wider
  select(f_concentration_group, lives_in, percentage) |>
  # the plot
  ggplot(aes(
    x = lives_in,
    label = percentage,
    y = percentage,
    fill = f_concentration_group
  )) +
  geom_col() +
  # adjust the scale
  scale_y_continuous(labels = label_percent(scale = 1)) +
  # change the colors
  scale_fill_brewer(palette = "Spectral") +
  # the labels
geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5), size = 3) +
  labs(subtitle = "Lithuania", 
       x = "Living Area group", 
       y = "", 
       fill = "F Concentration Group", 
       caption = "Note: Due to rounding, the sum of percentages may not equal 100.")
```

```{r}
latvia_Living_area_plot / lithuania_Living_area_plot
  
```

```{r}
ggsave(
  here("figures", "figure_Living_area.pdf"),
  width = 20, height = 20, units = "cm", 
  # res = 300, 
  device='pdf'
)
```

```{r}
rm(latvia_Living_area_plot, lithuania_Living_area_plot)
```

## Differences per living area

```{r}
df |> 
  select(lives_in, f_concentration_group) |> 
  gtsummary::tbl_summary(by = f_concentration_group) |> 
  add_p()
```

```{r}
chisq.test(table(df$lives_in, df$f_concentration_group)) 
```

```{r}
mosaicplot(table(df$f_concentration_group, df$lives_in), shade = T)
```

## Latvia per living area

```{r}

df_latvia |> 
  # filter(country == "Latvia") |> 
  select(lives_in, f_concentration_group) |> 
  gtsummary::tbl_summary(by = f_concentration_group) |> 
   modify_caption("**Latvia** (N = {N})") |> 
  add_p() 
 
```

```{r}
chisq.test(table(df_latvia$lives_in, df_latvia$f_concentration_group))
```

```{r}
mosaicplot(table(df_latvia$f_concentration_group, df_latvia$lives_in), shade = T)
```

## Lithuania per living area

```{r}

df_lithuania |> 
  # filter(country == "Latvia") |> 
  select(lives_in, f_concentration_group) |> 
  gtsummary::tbl_summary(by = f_concentration_group) |> 
   modify_caption("**Lithuania** (N = {N})") |> 
  add_p() 
 

# table(df_lithuania$lives_in, df_lithuania$f_concentration_group)
```

```{r}
chisq.test(table(df_lithuania$lives_in, df_lithuania$f_concentration_group))
```

```{r}
mosaicplot(table(df_lithuania$f_concentration_group, df_lithuania$lives_in), shade = T)
```

# F concentration analysis

```{r table concentration}
# janitor::tabyl(df$f_concentration) |> 
#   adorn_pct_formatting()
```

```{r}
df |> 
  ggplot(aes(x = f_concentration)) + 
  geom_histogram(bins = 10) + 
  scale_x_log10()
```

### Table f concentration by age with means, etc

```{r table F by group age mean sd etc}
df |> 
  group_by(`f_concentration_group`) |> 
  summarise(n = n(), 
            mean = mean(`age`), 
            sd = sd(`age`), 
            min = min(`age`), 
            max = max(`age`)) |> 
  mutate_if(is.numeric, round, 1)|> 
  kable() |> 
  kable_classic(full_width = F)
```

```{r}
df$f_concentration_test <- factor(df$f_concentration, 
                             levels = c(
                              "NA",
                              "No fluoride", 
                              "<1000 ppm", 
                              "551 to 999 ppm", 
                              "1000 and 1500 ppm"))
```

### Table f concentration by age

```{r table age by concentration}
df |> 
  janitor::tabyl(age_group, `f_concentration`) |> 
  kable() |> 
  kable_classic(full_width = F)
```

### Table f concentration by Live in

```{r table live by concentration}
df |> 
  janitor::tabyl(`lives_in`, `f_concentration`) |> 
  janitor::adorn_totals(c("row", "col")) |>
  janitor::adorn_percentages("row") |> 
  janitor::adorn_pct_formatting(rounding = "half up", digits = 0) |>
  janitor::adorn_ns() |>
  janitor::adorn_title("combined") |> 
  kable() |> 
  kable_classic(full_width = F)
```

### Table f concentration by age group

```{r nice table age by concentration}
df |> 
  janitor::tabyl(age_group, `f_concentration`) |> 
  janitor::adorn_totals(c("row", "col")) |>
  janitor::adorn_percentages("row") |> 
  janitor::adorn_pct_formatting(rounding = "half up", digits = 0) |>
  janitor::adorn_ns() |>
  janitor::adorn_title("combined") |> 
  knitr::kable() |> kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r see names F free toothpastes }
# df |>  
#   filter(`f_concentration` == "0 - No fluoride") |> 
#   na.omit() |>    # Using "data", filter out all rows with NAs in aa 
#   group_by(`toothpaste`) |>          # Then, with the filtered data, group it by "bb"
#   summarise(Unique_Elements = n_distinct(`toothpaste`))  |>   # Now summarise with unique elements per group
#   kable() |> 
#   kable_classic(full_width = F)
```

# Number of different toothpastes

```{r}
df |> 
  mutate(toothpaste2 = fct_lump_prop(toothpaste, prop = .02)) |> 
  group_by(toothpaste2, country) |> 
  count() |> 
  pivot_wider(names_from = country, 
              values_from = n) |> 
  arrange(desc(Lithuania)) |> 
  arrange(desc(Latvia)) |> 
  rename(Toothpaste = toothpaste2) |> 
   kable() |> 
  kable_classic(full_width = F)
```

## Full list toothpastes by country

```{r check toothpastes names}
df |> 
  group_by(toothpaste, country) |> 
  summarise(count = n()) |> 
  arrange(desc(count)) |> 
  pivot_wider(names_from = country,
              values_from = count) |> 
  rename("Toothpaste Brand" = "toothpaste") |> 
  kable() |> 
  kable_classic(full_width = F)
  
```

```{r see names all toothpastes }
# df |>  
#   filter(!`f_concentration` == "0 - No fluoride") |> 
#   #na.omit() |>    # Using "data", filter out all rows with NAs in aa 
#   group_by(`toothpaste`) |>          # Then, with the filtered data, group it by "bb"
#   summarise(Unique_Elements = n_distinct(`toothpaste`))  |>   # Now summarise with unique elements per group
#   knitr::kable()
```

# Different ways of presenting the differences between groups (not for publication)

Check groups of F concentration by age

```{r F by age boxplot}

df |> 
   filter(f_concentration_group != "NA") |>
  ggplot(aes(x = `f_concentration_group`, y = `age`)) +
  
  geom_jitter(alpha = .05) +
  
  geom_boxplot(alpha = .8) +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "Age distribution per toothpaste F concentration type", 
       x = "Fluoride concentration", 
       y = "Age (years)") 

```

```{r plot porcentaje por edad}
df |>
  filter(f_concentration_group != "NA") |>
  group_by(age_group, f_concentration_group) |>
  summarise(count = n()) |>
  mutate(perc = count / sum(count)) |>
  ggplot(aes(x = age_group, y = perc * 100, fill = f_concentration_group)) +
  geom_bar(stat = "identity") +
  facet_grid(f_concentration_group ~ .) +
  theme_minimal() +
  labs(title = "Fluoride Concentration toothpaste by age group",
       y = "Percentage",
       x = "Age group",
       fill = "Fluoride concentration") +
  scale_fill_brewer(palette = "Spectral")

```

## Figure 1

```{r}
df |> 
  filter(f_concentration_group != "NA") |> 
  # filter(f_concentration_group != "Not available") |> 
  group_by(age_group, f_concentration_group) |> 
  summarise(count = n()) |> 
  mutate(perc=count/sum(count)) |> 
  ggplot(aes(x = age_group, y = perc*100, fill = f_concentration_group)) +
  geom_bar(stat="identity") + 
  facet_wrap(f_concentration_group ~ . ) + 
  theme_minimal() + 
  labs(title = "Fluoride Concentration toothpaste by age group", 
       y = "Percentage", 
       x = "Age group", 
       fill = "Fluoride concentration") + 
  # ylim(c(0, 70)) + 
  coord_flip() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("#E69F00", "#882E72", "#D55E00", "#F0E442", "#009E73")) + 
  scale_y_continuous(labels = label_percent(scale = 1))
  
```

```{r plot by city}
df |>
  filter(f_concentration_group != "NA") |> 
  group_by(lives_in, f_concentration_group) |>
  summarise(count = n()) |>
  mutate(perc = count / sum(count)) |>
  ggplot(aes(x = lives_in, y = perc * 100, fill = f_concentration_group)) +
  geom_bar(stat = "identity") +
  facet_wrap( ~ f_concentration_group) +
  theme_minimal() +
  labs(title = "Fluoride Concentration toothpaste by location",
       y = "Percentage",
       x = "Location",
       fill = "Fluoride concentration") +
  coord_flip() +
  scale_fill_brewer(palette = "Spectral")
```

```{r}
df |> 
  tabyl(age_group, f_concentration_group ) |> 
  adorn_percentages() |> 
  adorn_pct_formatting()
```

# Codebook

```{r}
# df |> 
#  select(-c(family_code, nr_of_family_members)) |> 
#  dataReporter::makeCodebook()
```
