---
title: "Toothpaste Latvia 20230208"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    fig_caption: yes
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
  word_document:
    toc: yes
---
# Packages
```{r}
pacman::p_load(tidyverse,
               kableExtra, 
               here, 
               janitor,
               MatchIt,
               scales, 
   
                table1, # contains the label function
               gtsummary, 

               ggpubr, 
               RColorBrewer, 


                leaflet,

               viridis)
```

# Dataset


```{r read csv, include=FALSE}
df <- read_csv("01_data.csv") 
```

```{r}
theme_set(theme_minimal())
```


```{r}
df <- janitor::clean_names(df)
```

# Data cleaning
```{r visdat}
# visdat::vis_dat(df)
```

```{r}
names(df)
```

Clean names and remove empty cols


Age groups
```{r create age groups}
df <- df %>% 
  mutate(age_group = case_when(
    `age` < 6 ~ "0 ~ 5", 
    `age` < 13 ~ "6 ~ 12", 
    `age` < 19 ~ "13 ~ 18", 
    TRUE ~ "> 18"
  ))


#  relevel age groups
df$age_group <- fct_relevel(df$age_group, "0 ~ 5" , "6 ~ 12", "13 ~ 18", "> 18" )

```

```{r}
janitor::tabyl(df$age_group) %>% janitor::adorn_pct_formatting() # # table age groups
```


IN age group change NA to adult

```{r}
df$`children_adult` <- df$`children_adult` %>% 
  replace_na("adult")
```

## Grouping fluoride concentrations
Create a new variable F_concentration_group

```{r F groups}
df <- df %>% 
  mutate(`f_concentration_group` = case_when(
    `f_concentration` < 1 ~ "No fluoride",
    `f_concentration` < 1000 ~ "<1000 ppm",
    `f_concentration` < 1400 ~ "1000-1399 ppm", 
    `f_concentration` < 1501 ~ "1400-1500 ppm", 
    TRUE ~ "NA"
  )) %>% 
  mutate(`f_concentration_group` = fct_relevel(f_concentration_group, "No fluoride", "<1000 ppm","1000-1399 ppm", "1400-1500 ppm" ))
```


# Separate datasets for each country 

```{r}
df_latvia <- df %>% 
  filter(country == "Latvia")
  
```

```{r}
df_lithuania <- df %>% 
  filter(country == "Lithuania")
  
```


# EDA
```{r check names}
names(df)
```
```{r summary df}
summary(df)
```
```{r summary LV}
summary(df_latvia)
```
```{r summary LT}
summary(df_lithuania)
```


Convert F to number
```{r Convert F to number}
df$`f_concentration` <- as.integer(df$`f_concentration`)
```

```{r Check distribution of F in toothpaste}
df %>% 
  ggplot(aes(x = `f_concentration`)) + 
  geom_histogram(bins = 5)
```

Convert F2 to number
```{r Convert F2 to number}
# df$`f_concentration_2` <- as.integer(df$`f_concentration_for_the_second_paste`)
```

```{r Check distribution of F2 in toothpaste}
# df %>% 
#  ggplot(aes(x = `f_concentration_for_the_second_paste`)) + 
#  geom_histogram(bins = 5)
```


#Table 1
```{r}
df$children_adult <-
  factor(df$children_adult,
         labels = c("Adult toothpaste",
                    "Child toothpaste"))

label(df$age) <- "Age"
label(df$age_group) <- "Age groups"
label(df$lives_in) <- "Lives In"
label(df$lives_in_specific) <- "Region"
label(df$f_concentration_group) <- "F concentration in toothpaste"
label(df$nr_of_family_members) <- "Number of family members"

```


```{r}
df %>% 
  select(age, age_group, lives_in, country) %>% 
  gtsummary::tbl_summary(by = country, missing = "no", 
                         statistic = list(all_continuous() ~ "{mean} ({sd})")) %>% 
  gtsummary::add_overall()
```

# Table of main results

```{r}
df %>% 
  select(f_concentration_group, children_adult, country) %>% 
  gtsummary::tbl_summary(by = country, missing = "no", 
                         statistic = list(all_continuous() ~ "{mean} ({sd})")) %>% 
  gtsummary::add_overall()
```


```{r Table 1_1}
df %>% 
  select(lives_in, f_concentration_group) %>% 
  gtsummary::tbl_summary(by = lives_in, missing = "no") %>% 
  gtsummary::add_overall()
```

## Table of main results for Latvia
```{r Table 1_2}
df %>% 
  filter(country == "Latvia") %>% 
  select(f_concentration_group, age_group) %>% 
  gtsummary::tbl_summary(by = age_group, missing = "no") %>% 
  gtsummary::add_overall()
```


```{r Table 1_3}
df %>% 
  filter(country == "Latvia") %>% 
  select(lives_in, f_concentration_group) %>% 
  gtsummary::tbl_summary(by = lives_in, missing = "no") %>% 
  gtsummary::add_overall()
```

## Table of main results for Lithuania

```{r}
df %>% 
  filter(country == "Lithuania") %>% 
  select(f_concentration_group, age_group) %>% 
  gtsummary::tbl_summary(by = age_group, missing = "no") %>% 
  gtsummary::add_overall()
```

```{r Table 1}
df %>% 
  filter(country == "Lithuania") %>% 
  select(lives_in, f_concentration_group) %>% 
  gtsummary::tbl_summary(by = lives_in, missing = "no") %>% 
  gtsummary::add_overall()
```


# Graph F concentration by age groups
Sergio, por favor, agrega valores en grafico

```{r}
my_palette <- brewer.pal(name="Spectral", n = 7)[0:9]
```


```{r}
df %>% 
  select(age_group, country, f_concentration_group) %>% 
  filter(f_concentration_group != "NA") %>% 
  
  # calculate the percentages
  group_by(country, f_concentration_group, age_group) %>% 
  count() %>% 
  # pivot_wider(names_from = f_concentration_group,
  #            values_from = n)

  group_by(country) %>% 
  mutate(perc = n / sum(n) * 100) %>% 
  mutate(perc = round(perc, 1)) %>% 
  # select(-n) %>% 
  # pivot_wider(names_from = f_concentration_group,
   #           values_from = perc)
  
 
   # make the plot
   ggplot(aes(x = age_group, 
             y = perc, 
             label = perc, 
             fill = f_concentration_group)) +
  geom_col(position="fill") +
  
  scale_fill_brewer(palette = "Spectral") +
  labs(title = "Use of F Toothpaste by Age", 
       fill = "Concentration",
       x = "Age group", 
       y = "", 
       caption = "Numbers on bars corresponds to the number of persons. Latvia N = 2 270, Lithuania = 2 815."
       ) + 
  #geom_text(
   # aes(label = n),
    #size = 2.5, 
    #position = position_fill(0.5)
  # ) +
  scale_y_continuous(labels = label_percent()) +
  # divide by country
  facet_grid( country ~ .) 
```

```{r}
ggsave(
  here("figures", "figure_age.pdf"),
  width = 20, height = 20, units = "cm", 
  # res = 300, 
  device='pdf'
)
```


## Differences per age group
```{r}
table(df$age_group, df$f_concentration_group)
```
```{r}
chisq.test(table(df$age_group, df$f_concentration_group))
```
```{r}
mosaicplot(table(df$f_concentration_group, df$age_group), shade = T)
```
## Latvia per age
```{r}

table(df_latvia$age_group, df_latvia$f_concentration_group)
```
```{r}
chisq.test(table(df_latvia$age_group, df_latvia$f_concentration_group))
```
```{r}
mosaicplot(table(df_latvia$f_concentration_group, df_latvia$age_group), shade = T)
```


## Lithuania per age
```{r}
  table(df_lithuania$age_group, df_lithuania$f_concentration_group)
```
```{r}
chisq.test(table(df_lithuania$age_group, df_lithuania$f_concentration_group))
```
```{r}
mosaicplot(table(df_lithuania$f_concentration_group, df_lithuania$age_group), shade = T)
```


# Graph F concentration by living area
Sergio, por favor, agrega valores en grafico

```{r}
df %>% 
  select(lives_in, country, f_concentration_group) %>% 
  filter(f_concentration_group != "NA")%>% 
  
    # calculate the percentages
  group_by(country, f_concentration_group, lives_in) %>% 
  count() %>% 
 

  group_by(country) %>% 
  mutate(perc = n / sum(n) * 100) %>% 
  mutate(perc = round(perc, 1)) %>% 
  
  # make the plot

   # make the plot
   ggplot(aes(x = lives_in, 
             y = perc, 
             label = perc, 
             fill = f_concentration_group)) +
  geom_col(position="fill") +
  
  scale_fill_brewer(palette = "Spectral") +
  labs(title = "Use of F Toothpaste by Area", 
       fill = "Concentration",
       x = "Area", 
       y = "", 
       caption = " Numbers on bars corresponds to the number of persons. Latvia N = 2 270, Lithuania = 2 815."
       ) + 
#  geom_text(
#    aes(label = perc),
#    size = 2.5, 
#    position = position_fill(0.5)
#  ) +
  scale_y_continuous(labels = label_percent()) +
  # divide by country
  facet_grid( country ~ .) 
```

```{r}
ggsave(
  here("figures", "figure_area.pdf"),
  width = 20, height = 20, units = "cm", 
  # res = 300, 
  device='pdf'
)
```


## Differences per living area

```{r}

table(df$lives_in, df$f_concentration_group)
```
```{r}
chisq.test(table(df$lives_in, df$f_concentration_group))
```
```{r}
mosaicplot(table(df$f_concentration_group, df$lives_in), shade = T)
```

## Latvia per living area

```{r}

table(df_latvia$lives_in, df_latvia$f_concentration_group)
```
```{r}
chisq.test(table(df_latvia$lives_in, df_latvia$f_concentration_group))
```
```{r}
mosaicplot(table(df_latvia$f_concentration_group, df_latvia$lives_in), shade = T)
```

## Lithuania per living area

```{r}

table(df_lithuania$lives_in, df_lithuania$f_concentration_group)
```
```{r}
chisq.test(table(df_lithuania$lives_in, df_lithuania$f_concentration_group))
```
```{r}
mosaicplot(table(df_lithuania$f_concentration_group, df_lithuania$lives_in), shade = T)
```

# F concentration analysis

```{r table concentration}
janitor::tabyl(df$f_concentration) %>% 
  adorn_pct_formatting()
```

```{r}
df %>% 
  ggplot(aes(x = f_concentration)) + 
  geom_histogram(bins = 10)
```

Table f concentration by age with means, etc
```{r table F by group age mean sd etc}
df %>% 
  group_by(`f_concentration_group`) %>% 
  summarise(n = n(), 
            mean = mean(`age`), 
            sd = sd(`age`), 
            min = min(`age`), 
            max = max(`age`)) %>% 
  mutate_if(is.numeric, round, 1)
```


```{r}
df$f_concentration_test <- factor(df$f_concentration, 
                             levels = c(
                              "NA",
                              "No fluoride", 
                              "<1000 ppm", 
                              "551 to 999 ppm", 
                              "1000 and 1500 ppm"))
```



```{r table age by concentration}
df %>% 
  janitor::tabyl(age_group, `f_concentration`) %>% 
  knitr::kable()
```


```{r table live by concentration}
df %>% 
  janitor::tabyl(`lives_in`, `f_concentration`) %>% 
  janitor::adorn_totals(c("row", "col")) %>%
  janitor::adorn_percentages("row") %>% 
  janitor::adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  janitor::adorn_ns() %>%
  janitor::adorn_title("combined") %>%
  knitr::kable()
```


```{r nice table age by concentration}
df %>% 
  janitor::tabyl(age_group, `f_concentration`) %>% 
  janitor::adorn_totals(c("row", "col")) %>%
  janitor::adorn_percentages("row") %>% 
  janitor::adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  janitor::adorn_ns() %>%
  janitor::adorn_title("combined") %>%
  knitr::kable() %>% kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```



```{r see names F free toothpastes }
df %>%  
  filter(`f_concentration` == "0 - No fluoride") %>% 
  na.omit() %>%    # Using "data", filter out all rows with NAs in aa 
  group_by(`toothpaste`) %>%          # Then, with the filtered data, group it by "bb"
  summarise(Unique_Elements = n_distinct(`toothpaste`))  %>%   # Now summarise with unique elements per group
  knitr::kable()
```



# Number of different toothpastes


```{r}
df %>% 
  mutate(toothpaste2 = fct_lump_prop(toothpaste, prop = .02)) %>% 
  group_by(toothpaste2, country) %>% 
  count() %>% 
  pivot_wider(names_from = country, 
              values_from = n) %>% 
  arrange(desc(Lithuania)) %>% 
  arrange(desc(Latvia))
```



```{r check toothpastes names}
df %>% 
  group_by(toothpaste, country) %>% 
  summarise(count = n())
```



```{r see names all toothpastes }
df %>%  
  filter(!`f_concentration` == "0 - No fluoride") %>% 
  #na.omit() %>%    # Using "data", filter out all rows with NAs in aa 
  group_by(`toothpaste`) %>%          # Then, with the filtered data, group it by "bb"
  summarise(Unique_Elements = n_distinct(`toothpaste`))  %>%   # Now summarise with unique elements per group
  knitr::kable()
```


# Different ways of presenting the differences between groups (not for publication)

Check groups of F concentration by age
```{r F by age boxplot}

df %>% 
   filter(f_concentration_group != "NA") %>%
  ggplot(aes(x = `f_concentration_group`, y = `age`)) +
  
  geom_jitter(alpha = .05) +
  
  geom_boxplot(alpha = .8) +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "Age distribution per toothpaste F concentration type", 
       x = "Fluoride concentration", 
       y = "Age (years)") 

```


```{r plot porcentaje por edad}
df %>%
  filter(f_concentration_group != "NA") %>%
  group_by(age_group, f_concentration_group) %>%
  summarise(count = n()) %>%
  mutate(perc = count / sum(count)) %>%
  ggplot(aes(x = age_group, y = perc * 100, fill = f_concentration_group)) +
  geom_bar(stat = "identity") +
  facet_grid(f_concentration_group ~ .) +
  theme_minimal() +
  labs(title = "Fluoride Concentration toothpaste by age group",
       y = "Percentage",
       x = "Age group",
       fill = "Fluoride concentration") +
  scale_fill_brewer(palette = "Spectral")

```
```{r}
df %>% 
  filter(f_concentration_group != "NA") %>% 
  group_by(age_group, f_concentration_group) %>% 
  summarise(count = n()) %>% 
  mutate(perc=count/sum(count)) %>% 
  ggplot(aes(x = age_group, y = perc*100, fill = f_concentration_group)) +
  geom_bar(stat="identity") + 
  facet_wrap(~f_concentration_group) + 
  theme_minimal() + 
  labs(title = "Fluoride Concentration toothpaste by age group", 
       y = "Percentage", 
       x = "Age group", 
       fill = "Fluoride concentration") + 
  coord_flip() +
  scale_fill_brewer(palette = "Spectral")
```



```{r plot by city}
df %>%
  filter(f_concentration_group != "NA") %>% 
  group_by(lives_in, f_concentration_group) %>%
  summarise(count = n()) %>%
  mutate(perc = count / sum(count)) %>%
  ggplot(aes(x = lives_in, y = perc * 100, fill = f_concentration_group)) +
  geom_bar(stat = "identity") +
  facet_wrap( ~ f_concentration_group) +
  theme_minimal() +
  labs(title = "Fluoride Concentration toothpaste by location",
       y = "Percentage",
       x = "Location",
       fill = "Fluoride concentration") +
  coord_flip() +
  scale_fill_brewer(palette = "Spectral")
```

# Codebook
```{r}
# df %>% 
#  select(-c(family_code, nr_of_family_members)) %>% 
#  dataReporter::makeCodebook()
```






